### 2022-11-22 paper-thesis/chapters/21-adding-inference.tex
Finish one chapter
Corrections

\chapter{Term Inference}

@@

There are a lot of different algorithms for inference\cite{typeinference}, some algorithms can solve more inferences than others. However, we would like to avoid implementing an algorithm at all, instead using Statix's built-in first-order unification to do the type inference for us. Implementing an inference algorithm in Statix is theoretically possible but this would be a lot of ugly code, and the goal is to use Statix in a way that is clean and declarative, not to do optimal inference.

@@

There are some situations in which this approach fails, but in most real-world scenarios it works perfectly. For example, it can infer both programs in the introduction of this chapter, but it fails to infer the following program:
\begin{lstlisting}
let f = _;
\x : Type.
\g: (_: (f x) -> Bool).
g true
\end{lstlisting}
We know that \verb|f| is a function from \verb|Type -> Type|, but it fails to infer the value of \verb|f|. Because of the way that \verb|g| is used, the type checker asserts that $\beq{f x}{Bool}$. Since x is declared as a function argument and it is completely free, this means that for any \verb|x|, \verb|f x = Bool|. But the rule above is not powerful enough to derive all of this, so it fails.

\section{Using Statix's first-order unification for inference}
\label{statix-inference}

First, we introduce a new constructor, \verb|Infer : Expr -> Expr|. This will be used to store metavariables which we want to infer. This is a hack we use, ideally we would like to query statix to see if a metavariable is initialized, but as discussed in section \ref{strength-inference} this is not possible.

>>>>

\chapter{Term Inference}
\label{ch:inference}

@@

There are a lot of different algorithms for inference\cite{typeinference}, some algorithms can solve more inferences than others. However, we would like to avoid implementing an algorithm at all, instead using Statix' built-in first-order unification to do the type inference for us. Implementing an inference algorithm in Statix is theoretically possible but this would be a lot of ugly code (since Statix is not a general-purpose programming language), and the goal is to use Statix in a way that is clean and declarative, not to do optimal inference.

@@

There are some situations in which this approach fails, but in most real-world scenarios it works perfectly. For example, it can infer both programs in the introduction of this chapter, but it fails to infer the following program:

\begin{lstlisting}
let f = _;
\x : Type.
\g: (_: (f x) -> Bool).
g true
\end{lstlisting}
We know that \verb|f| is a function from \verb|Type -> Type|, but it fails to infer the value of \verb|f|. Because of the way that \verb|g| is used, the type checker asserts that $\beq{f x}{Bool}$. Since x is declared as a function argument and it is completely free, this means that for any \verb|x|, \verb|f x = Bool|. But the rule above is not powerful enough to derive all of this, so it fails.

\section{Using Statix' first-order unification for inference}
\label{statix-inference}

First, we introduce a new constructor, \verb|Infer : Expr -> Expr|. This will be used to store metavariables which we want to infer. This is a hack we use, ideally we would like to query Statix to see if a metavariable is initialized, but as discussed in section \ref{strength-inference} this is not possible.
